\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=2.5cm}

\title{\textbf{Scalable Document Reading with LLM-Augmented Navigation}}
\author{A. R. Ciment \and D. A. Cask\\
        Department of Computer Science\\
        University of Imaginary Sciences\\
        Dreamland}
\date{}

\begin{document}
\maketitle

\begin{abstract}
We present a lightweight architecture for browser-based scientific PDF readers that
combines high-fidelity canvas rendering with large language model assistance. Our
system renders pages using PDF.js, synchronizes a text-layer for selectable content,
and exposes an AI side panel for contextual summarization and question answering. We
evaluate rendering latency across document lengths up to one hundred pages and show
that the proposed approach maintains interactive frame rates. A user study with
thirty-two participants indicates that integrated navigation and conversational
explanations reduce the time required to locate key sections by approximately forty
percent.
\end{abstract}

\section{Introduction}

Reading long technical documents in the browser remains a frustrating experience.
Users frequently switch between a PDF viewer, a note-taking application, and a
separate chat interface to ask questions about the content. Existing web-based
readers either render documents faithfully but offer no assistance, or they extract
plain text and lose the original typography and figures.

We introduce \emph{Cimedaca Reader}, a single-page application built on SvelteKit that
marries faithful page rendering with a conversational AI companion. The reader loads
the Portable Document Format binary in the browser, renders each page to a canvas,
overlays a transparent text layer for selection and search, and displays an AI panel
on the right side of the screen.

The main contributions of this paper are:
\begin{itemize}
    \item a reactive rendering pipeline that couples PDF.js with Svelte~5 runes,
    \item a local-only text extraction strategy that preserves user privacy,
    \item and an empirical evaluation showing interactive performance on documents
    up to one hundred pages.
\end{itemize}

\section{System Architecture}

The application is organized into three panes. The navigation pane on the left lists
document sections and page thumbnails. The content pane in the center displays the
rendered pages. The AI pane on the right hosts a streaming chat interface backed by
an OpenAI-compatible endpoint.

\subsection{Rendering Pipeline}

The rendering pipeline begins when the user selects a file. The binary is stored in
memory as an \texttt{ArrayBuffer} and passed to \texttt{pdfjs-dist}. The library parses
the file, creates a document proxy, and exposes the number of pages. For each page
that enters the viewport, we create a canvas element, request the page proxy, and call
\texttt{render} with a scale derived from the current zoom level.

To support text selection, we render a transparent text layer over each canvas. The
text layer is populated with absolutely positioned \texttt{span} elements whose locations
and font sizes are computed from the page viewport. This preserves the visual fidelity
of the canvas while allowing the browser to handle selection, copying, and screen readers.

The zoom level $s$ is constrained to a minimum $s_{\min}=0.25$ and a maximum
$s_{\max}=3.0$. When the user triggers a zoom-in event, the new scale is computed as
\begin{equation}
    s' = \min(s \cdot \alpha, s_{\max}),
\end{equation}
where $\alpha = 1.1$ is the zoom step factor. Conversely, zoom-out uses
$\alpha^{-1}$ and clamps to $s_{\min}$.

\subsection{Zoom and Scroll}

Zoom is controlled through a reactive scale value in a Svelte rune store. When the user
presses \texttt{Control} and scrolls the mouse wheel, the scale is multiplied by a constant
factor. A fit-to-width button computes the largest scale that fits the available client
width for the first page. Scrolling is handled by a single overflow container that stacks
pages vertically.

The user can also jump directly to a page by clicking an outline entry or a thumbnail. The
target page element is scrolled into view with smooth behavior. Formally, the visible page
$v$ at scroll offset $y$ is
\begin{equation}
    v = \arg\max_{p} \, \bigl( \min(y_p + h_p, y + H) - \max(y_p, y) \bigr),
\end{equation}
where $y_p$ and $h_p$ are the offset and height of page $p$, and $H$ is the viewport height.

\section{Language Model Integration}

The AI pane submits the current document context together with the user's prompt. We first
extract the visible page text from the active text layers. This text is concatenated,
truncated to fit the model context window, and sent as a system message. The user's question
is appended and streamed back token by token.

We deliberately avoid uploading the raw binary to any external service. All text extraction
happens locally in the browser, so the document contents never leave the user's device unless
the user explicitly pastes them into the chat.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
    box/.style={draw, rounded corners, minimum width=3cm, minimum height=0.9cm, align=center}
]
\node[box, fill=gray!15] (pdf) at (0,0) {PDF file};
\node[box] (parse) at (4.2,0) {PDF.js parser};
\node[box] (render) at (8.4,0) {Canvas renderer};
\node[box] (text) at (8.4,-2) {Text layer};
\node[box] (llm) at (4.2,-2) {Local LLM client};

\draw[->, thick] (pdf) -- (parse);
\draw[->, thick] (parse) -- (render);
\draw[->, thick] (parse) -- ++(0,-1.2) -| (text);
\draw[->, thick, dashed] (text) -- (llm);
\end{tikzpicture}
\caption{Data flow inside the reader. Solid lines represent rendering paths; the dashed line
shows how extracted text is routed to the local language model client.}
\label{fig:architecture}
\end{figure}

\section{Evaluation}

We measured rendering latency on a synthetic corpus of ten documents ranging from five to one
hundred pages. Each document was generated with a mixture of paragraphs, equations, tables,
and vector figures. Tests were run on a laptop with a four-core processor and eight gigabytes
of memory.

\begin{table}[htbp]
\centering
\caption{Rendering latency and scroll performance across document lengths.}
\label{tab:latency}
\begin{tabular}{@{}rrrr@{}}
\toprule
Pages & First paint (ms) & Layout complete (ms) & Scroll fps \\
\midrule
5  & 120 & 180 & 60 \\
20 & 145 & 260 & 60 \\
50 & 210 & 580 & 55 \\
100 & 380 & 1120 & 48 \\
\bottomrule
\end{tabular}
\end{table}

The results in Table~\ref{tab:latency} show that first paint remains below four hundred
milliseconds even for the longest document. Layout completion time grows linearly with page
count because every canvas is created upfront. Future work will investigate virtualized
rendering to reduce initial layout time.

We also conducted a small user study with thirty-two graduate students in computer science.
Each participant was asked to locate three specific claims in a twenty-page paper. Half of
the participants used a standard PDF viewer, while the other half used Cimedaca Reader with
the AI pane enabled. The AI-assisted group completed the task in an average of four minutes
and twelve seconds, compared to seven minutes and five seconds for the control group.

\section{Related Work}

Browser-based PDF rendering has been dominated by Mozilla's PDF.js since its release in 2011.
PDF.js parses PDF byte streams, interprets page descriptions, and renders them to HTML5 canvas
elements. Many readers wrap PDF.js in different frameworks but do not integrate language models.

Recent AI assistants for documents either require uploading files to a server or convert PDFs
to Markdown before processing. Our approach differs by keeping the rendering pipeline and the
language model client fully local. This design aligns with the local-first software principles
described by Kapoor, in which users retain ownership of their data.

\section{Conclusion}

We described a reader architecture that renders PDFs faithfully in the browser while providing
an AI assistant for navigation and summarization. The system achieves interactive frame rates on
moderately long documents and keeps user data private by performing text extraction locally.
Source code and benchmarks are available at \url{https://example.org/cimedaca}.

\begin{thebibliography}{9}
\bibitem{jones2021}
B. Jones and L. Smith.
Adaptive layouts for responsive PDF readers.
In \emph{Proceedings of the Web Systems Conference}, pages 45--58, 2021.

\bibitem{kapoor2023}
A. Kapoor.
Local first software: you own your data.
\emph{The Local-First Journal}, 2023.

\bibitem{wang2022}
J. Wang, K. Patel, and M. Chen.
Evaluating text extraction quality from academic PDFs.
\emph{Information Retrieval Letters}, 19(3):210--225, 2022.
\end{thebibliography}

\end{document}
